---
title: "Documentation for the SUNFLO crop model"
output: rmarkdown::tufte_handout
bibliography: ./files/bibliography.bib
csl: ./style/elsevier.csl
---

```{r header, message=FALSE, warning=FALSE, echo=FALSE}
# TODO : made small parameters table linked subsets of the complete parameter table

library(ggplot2)
library(reshape2)
library(tidyr)
library(dplyr)
library(XLConnect)
library(knitr)

# path (! wd is changed in next chunk)
opts_knit$set(root.dir = normalizePath('../../'))

# options
fig_dpi = 100
```

```{r data, message=FALSE, warning=FALSE, echo=FALSE}
# data
db <- loadWorkbook(file="inst/doc/files/parameterization.xlsx") 

sunflo <- list(
  climate = readWorksheet(db, sheet="climate"),
  parameters = readWorksheet(db, sheet="parameters"),
  outputs = readWorksheet(db, sheet="outputs"),
  indicators = readWorksheet(db, sheet="indicators")
)
sunflo$parameters <- mutate(sunflo$parameters, value=as.numeric(gsub(",",".",value)))

```

\begin{figure*}
  \includegraphics[scale=2]{files/structure_nodes.pdf}
\end{figure*}

\newpage

\footnotesize
\tableofcontents

\newpage


```{r structure, echo=FALSE, eval=FALSE}
# Build model graph from R
graph_ext <- "png"
graph_nodes <- "-Nlabel= -Nshape=circle -Gdpi=150 -Earrowsize=2 -Epenwidth=4 -Estyle=dotted"

# Simple graph : replace nodes names by points
call <- paste0("dot -T ",graph_ext," -o inst/doc/files/structure_nodes.",graph_ext," ./files/structure_nodes.dot")
system(call)
# system(paste(call, graph_nodes))

# Complete graph
call <- paste0("dot -T ",graph_ext," -o inst/doc/files/structure.",graph_ext," ./files/structure.dot")
system(call)

```


# Inputs
```{r InputsParameters, echo=FALSE, results='asis'}
s <- sunflo$parameter %>%
  filter(count==1) %>%
  select(category, label.doc) %>%
  group_by(category) %>%
  summarise(n=n())

n_total <- sum(s$n)
```

## Climate
```{r InputsClimate, echo=FALSE, results='asis'}
# data
s <- sunflo$climate %>% select(label=label.doc, description=label.en, unit)
# table
kable(s, padding=0)
```

## Soil
```{r InputsSoil, echo=FALSE, results='asis'}
# data
s <- sunflo$parameters %>%
  filter(category=="Pedoclimat", template.default==1, name.vle!="datas_file") %>%
  select(label=label.doc, description=label.en, value, unit, reference)
# table
kable(s, padding=0)
```

## Management
```{r InputsManagement, echo=FALSE, results='asis'}
# data
s <- sunflo$parameters %>%
  filter(category=="Conduite", template.default==1) %>%
  select(label=label.doc, description=label.en, value, unit, reference)
# table
kable(s, padding=0)
```

## Species
```{r InputsSpecies, echo=FALSE, results='asis'}
```


## Cultivar
```{r InputsCultivar, echo=FALSE, results='asis'}
# data
s <- sunflo$parameters %>%
  filter(category=="Variete", template.default==1) %>%
  select(label=label.doc, description=label.en, value, unit, reference)
# table
kable(s, padding=0)
```


\newpage

# Phenology
```{r InputsPhenology, echo=FALSE, results='asis'}
# data
s <- sunflo$parameters %>% 
  filter(module=="Phenology") %>%
  select(label=label.doc, description=label.en, value, unit, reference)

# table
kable(s, padding=0)
```

## Emergence
$Emergence = Germination + ElongationRate * SowingDepth$

with:

* $Germination = 86$, Thermal time for germination (°C.d);
* $ElongationRate = 1.19$, Hypocotyl elongation rate (°Cd/mm)
* $SowingDepth = 30$, Sowing depth (mm).

## ThermalTime
$$
  ThermalTime_d=
	\left\{ 
	\begin{array}{ll}
	\int_0^d (T_m - T_b) * (1 + WaterStressPhenology)  & \textrm{if $T_m > T_b$} \\  
	0 & \textrm{else}
	\end{array} 
	\right.
$$

with: 

* $T_m$, Daily mean air temperature (°C); 
* $T_b = 4.8$, Basal temperature (°C) [@Granier1998]; 
* $ThermalStressPhenology$, Water stress effect on plant heating

## PhenoStages
```{r PhenoStages, echo=FALSE, fig.width=8, fig.height=2.5, fig.fullwidth=TRUE}
# Data
parameters <- c("ThermalTimeVegetative", "ThermalTimeFlowering",
                "ThermalTimeSenescence", "ThermalTimeMaturity",
                "Germination")

list_phase <- c("Vegetative","Floral Initiation","Flowering","Grain Filling","Germination","Elongation")

s <- sunflo$parameters %>%
  filter(module=="Phenology", label.doc %in% parameters) %>%
  select(label=label.doc, value)

# add emergence state
s <- rbind(s, data.frame(label="Emergence", value=0))
# change germination value according to emergence
s[s$label=="Germination",]$value <- -200+ s[s$label=="Germination",]$value
# compute start phase for plotting
s <- s %>% mutate(
  phase=factor(list_phase, levels=list_phase),
  start=c(0, 482, 836, 1083, -200, -200+86.2),
  end=value
)

# plot
p <- ggplot(s) +
  geom_segment(aes(x=start, xend=end, y=0, yend=0, color=phase), size=5) +
  geom_text(aes(x=value, y=0, label=label), vjust=0.5, size=3, angle=90, hjust=-0.1) +
  theme_bw() + labs(x="ThermalTime", y="") + ylim(0,0.5) +
  scale_color_manual(values = terrain.colors(6)) +
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank())

print(p)
# ggsave(p, file="./figure/PhenoStage.png", height=2.5, width=8, units="in", dpi=fig_dpi)
```

In the software model, phenostages are computed as integers $\in [0,7]$ corresponding to duration between key stages:

* **0**, begining of simulation - sowing (bare soil)
* **1**, sowing (A0) - emergence (A2)
* **2**, emergence (A2) - floral initiation (E1)
* **3**, floral initiation (E1) - flowering (F1)
* **4**, flowering (F1) - onset of senescence (M0)
* **5**, onset of senescence (M0) - maturity (M3)
* **6**, maturity (M3) - harvest
* **7**, harvest - end of simulation (bare soil)

\newpage

# LeafArea
```{r InputsLeafArea, echo=FALSE, results='asis'}
# data
s <- sunflo$parameters %>% 
  filter(module=="LeafArea") %>%
  select(label=label.doc, description=label.en, value, unit, reference)

# table
kable(s, padding=0)
```

\vspace{2cm}

## LeafInitiationTime, LeafExpansionTime, LeafSenescenceTime
```{r LeafInitiationTime, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
LeafInitiationTime <- function(i, phy_1=71.43, phy_7=16.34, i_max=30, a=420){
  ifelse(
    i <= 6, i*phy_1,
      ifelse(
        i <= i_max,
        (i-5) * phy_7 + a, NA
      )
  )
}

LeafExpansionDuration <- function(i, TLN=30, LLH=17, y0=851.3, a=153, b=0.78){
  a + y0 * exp(-((i - LLH)^2) / ((b * TLN)^2))
}

# Data
i <- 1:30

d <- data_frame(x=i) %>% 
  mutate(
    Initiation = LeafInitiationTime(x),
    Expansion = Initiation + 1/0.01379,
    Senescence = Expansion + LeafExpansionDuration(x)
  ) %>%
  gather(Process, y, -x)


# Graph
p <- ggplot(d, aes(x=y, y=x, color=Process)) + geom_path() +
  theme_bw() + labs(x="ThermalTime", y="Node") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)
```

$$
  LeafInitiationTime_i = \left\{ 
  \begin{array}{ll}
	i * Phyllotherm_1 & \textrm{if $i \leq 6$} \\  
	(i-5) * Phyllotherm_7 + a & \textrm{if $i \leq LeafNumber$}
	\end{array} \right.
$$

with:

* $Phyllotherm_1=76.43$ (°C.d)
* $Phyllotherm_7=16.34$ (°C.d)
* $a=400$ (°C.d)

$$LeafExpansionTime_i=LeafInitiation_i + 1/a$$

with $a=0.01379$.


## LeafExpansionDuration
```{r LeafExpansionDuration, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Data
i <- 1:30

d <- data_frame(x=i) %>%
  mutate(LeafExpansionDuration = LeafExpansionDuration(x)) %>%
  gather(Process, y, -x)

# Graph
p <- ggplot(d, aes(x, y, color=Process)) + geom_path() +
  theme_bw() + labs(x="Node", y="LeafExpansionDuration") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)

```

$$LeafExpansionDuration_i = a + PotentialLeafDuration * exp^{\frac{-(i-PotentialLeafProfile)^2}{(c * PotentialLeafNumber)^2}}$$

with: 

* $PotentialLeafDuration=851.3$ (°C.d)
* $a=153$ (°C.d)
* $b=0.78$

$$LeafSenescenceTime_i = LeafExpansionTime_i + LeafExpansionDuration_i$$


## PotentialLeafArea
```{r PotentialLeafArea, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
PotentialLeafArea <- function(i, LLS=448, LLH=17, a=-2.05, b=0.049) {
  LLS * exp(a*((i-LLH)/(LLH-1))^2 + b*((i-LLH)/(LLH-1))^3) 
}

# Data
i <- 1:30

d <- data_frame(x=i) %>%
  mutate(PotentialLeafArea = PotentialLeafArea(x)) %>%
  gather(Process, y, -x)

# Graph
p <- ggplot(d, aes(x, y, color=Process)) + geom_path() +
  theme_bw() + labs(x="Node", y="PotentialLeafArea") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)

```

$$
PotentialLeafArea_i = PotentialLeafSize * exp^{a * (\frac{i-PotentialLeafProfile}{PotentialLeafProfile-1})^2 + b *
(\frac{i-PotentialLeafProfile}{PotentialLeafProfile-1})^3}
$$

with: 
* $a=-2.05$ and $b=0.049$, shape parameters
* $PotentialLeafSize$ (cm2) and $PotentialLeafProfile$ (node), genotypic parameters.


## LeafGrowthRate, LeafSenescenceRate, LeafArea
```{r LeafGrowthRate, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
LeafGrowthRate <- function(t, te, Ae, Teff, Ke=0.01379){
  Teff * (Ae * Ke) * exp(-Ke * (t - te)) / (1 + exp(-Ke * (t - te)))^2
}

# Data
t <- 0:2000

d <- data_frame(x=t) %>% 
  mutate(
    Expansion = LeafGrowthRate(x, te=500, Ae=300, Teff=(25-4.8)),
    Senescence = LeafGrowthRate(x, te=1500, Ae=300, Teff=(25-4.8))
  ) %>%
  gather(Process, y, -x)

# Graph
p <- ggplot(d, aes(x, y, color=Process)) + geom_path() +
  theme_bw() + labs(x="ThermalTime (°Cd)", y="GrowthRate (cm2/°Cd)") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)
# ggsave(p, file="inst/doc/files/figures/LeafGrowthRate.pdf", height=3, width=3, units="in")
```

$$
\begin{split}
LeafGrowthRate_i = (T_m - T_b) * PotentialLeafArea_i * a * \\
\frac{exp^{-a(ThermalTime-LeafExpansionTime_i)}}{(1+exp^{-a(ThermalTime-LeafExpansionTime_i)})^2}
\end{split}
$$

$$
\begin{split}
LeafSenescenceRate_i = (T_m - T_b) * LeafArea_i * a * \\
\frac{exp^{-a(ThermalTime-LeafSenescenceTime_i)}}{(1+exp^{-a(ThermalTime-LeafSenescenceTime_i)})^2}
\end{split}
$$

with:

* $T_m=25$, mean air temperature (°C)
* $T_b=4.8$, base temperature (°C)
* $a=0.01379$

The illustration uses $i=10$ as values for $PotentialLeafArea_i$, $LeafExpansionTime_i$ and $LeafSenescenceTime_i$

$$LeafArea_i=\int LeafGrowthRate_i - \int LeafSenescenceRate_i$$


\newpage

# Light interception

## LAI
$$LAI = SowingDensity * \sum_{i=1}^{LeafNumber} LeafArea_i$$

## RIE
$$RIE = 1 - exp^{(-ExtinctionCoefficient * LAI)}$$

```{r RIE, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
RIE <- function(LAI, k) {1 - exp(-k * LAI)}

# Graph
lai <- seq(0,5, by=0.1)

d <- data_frame(x=lai) %>%
  mutate(
    low = RIE(x, k=0.7),
    high = RIE(x, k=0.9)
  ) %>%
  gather(k, y, -x)

# Graph
p <- ggplot(d, aes(x, y, color=k)) + geom_path() +
  theme_bw() + labs(x="Leaf Area Index (LAI)", y="RIE") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)
#ggsave(p, file="inst/doc/files/figures/RIE.pdf", height=3, width=3, units="in")
```


\newpage

# Biomass production

## RUE


$$
PotentialRUE = \left\{ 
	\begin{array}{ll}
	r_0 & \textrm{if $ThermalTime < 300 $} \\  
	r_0 + 2 * \frac{ThermalTime-300}{ThermalTimeFlowering-300} & \textrm{if $300 < ThermalTime < ThermalTimeFlowering$}\\
	r_{max} & \textrm{if $ThermalTimeFlowering < ThermalTime < ThermalTimeSenescence$}\\
	a * exp^{b * (1-\frac{ThermalTime - ThermalTimeMaturity}{ThermalTimeMaturity - ThermalTimeSenescence})} & \textrm{if $ThermalTimeSenescence < ThermalTime < ThermalTimeMaturity$}\\
  0 & \textrm{else}
	\end{array}
  \right.
$$

with:

* $r_0=1$, vegetative RUE
* $r_{max}=3$, maximum RUE
* $a=0.015$, final RUE
* $b=4.5$, slope of RUE decrease in grain filling stage

```{r RUE, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
PotentialRUE <- function(
  t,
  r0=1,
  rmax=3,
  ThermalTimeFlowering = 836,
  ThermalTimeSenescence = 1083,
  ThermalTimeMaturity = 1673,
  a = 0.015,
  b = 4.5
  ) {
  ifelse(
    t < 300, r0,
    ifelse(
      t >= 300 & t < ThermalTimeFlowering, r0 + 2*(t-300)/(ThermalTimeFlowering-300),
      ifelse(
        t >= ThermalTimeFlowering & t < ThermalTimeSenescence-100, rmax,
        ifelse(
          t >= ThermalTimeSenescence-100 & t <  ThermalTimeMaturity,
          a * exp(b*(1-((t-ThermalTimeSenescence)/(ThermalTimeMaturity-ThermalTimeSenescence)))), 0
        )
      )
    )
  )
}

# Graph
t <- 0:2000

d <- data_frame(x=t) %>% 
  mutate(PotentialRUE=PotentialRUE(x)) %>%
  gather(Process, y, -x)

p <- ggplot(d, aes(x, y, color=Process)) + geom_path() +
  theme_bw() + labs(x="Thermal Time (°C.d)", y="PotentialRUE") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)
#ggsave(p, file="inst/doc/files/figures/PotentialRUE.pdf", height=3, width=3, units="in")
```

\vspace{2cm}

## CropBiomass [@Monteith1977]
$$dCropBiomass = Radiation * 0.48 * RIE * RUE * dt$$

## CropPerformance


\newpage

# Thermal stress
## ThermalStressRUE [@Villalobos1996]
```{r ThermalStressRUE, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
ThermalStressRUE <- function(tm, tb=4.8, tol=20, tou=28, tc=37) {
  ifelse(
    tm > tb & tm < tol, tm * (1/(tol - tb)) - (tb/(tol - tb)),
    ifelse(
      tm >= tol & tm <= tou, 1,
      ifelse(
        tm > tou & tm < tc, tm * (1/(tou - tc)) - (tc/(tou-tc)),
        ifelse(
          tm <= tb | tm >= tc, 0, NA
        )
      )
    )
  )
}

# Graph
t <- seq(0,45,by=0.2)

d <- data_frame(x=t) %>% 
  mutate(ThermalStressRUE=ThermalStressRUE(x)) %>%
  gather(Process, y, -x)

p <- ggplot(d, aes(x, y, color=Process)) + geom_path() +
  theme_bw() + labs(x="Mean Air Temperature (°C)", y="ThermalStressRUE") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)
# ggsave(p, file="./figure/ThermalStressRUE.png", height=3.5, width=4, units="in", dpi=fig_dpi)
```


$$
ThermalStressRUE = \left\{ 
  \begin{array}{ll}
  T_m * \frac{1}{T_{ol} - T_b} - \frac{T_b}{T_{ol} - T_b} & \textrm{if $T_b < T_m < T_{ol}$} \\  
  1 & \textrm{if $T_{ol} < T_m < T_{ou}$} \\
  T_m * \frac{1}{T_{ou} - tc} - \frac{tc}{T_{ou} - tc} & \textrm{if $T_{ou} < T_m < tc$} \\
  0 & \textrm{else}
	\end{array}
  \right.
$$

with:

* $T_b = 4.8$, base temperature (°C)
* $T_{ol} = 20$, optimal lower temperature (°C)
* $T_{ou} = 28$, optimal upper temperature (°C)
* $T_c = 37$, critical temperature (°C)



## ThermalStressMineralization [@Vale2007]
```{r ThermalStressMineralization, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
ThermalStressMineralization <- function(tm, tb=15, tc=36) {
  tc/(1 + (tc - 1) * exp(-0.119 * (tm - tb)))
}

## Graph
t <- seq(0,45,by=0.2)

d <- data_frame(x=t) %>% 
  mutate(ThermalStressMineralization=ThermalStressMineralization(x)) %>%
  gather(Process, y, -x)

p <- ggplot(d, aes(x, y, color=Process)) + geom_path() +
  theme_bw() + labs(x="Mean Air Temperature (°C)", y="ThermalStressMineralization") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)

# ggsave(p, file="./figure/ThermalStressMineralization.png", height=3.5, width=4, units="in", dpi=fig_dpi)
```

$$ThermalStressMineralization = \frac{T_c}{1 + (T_c - 1) * exp^{(-0.119 * (T_m - T_b))}}$$

with: 

* $T_b = 15$, base temperature (°C)
* $T_c = 36$, critical temperature (°C)

## ThermalStressAllocation

\newpage

# Water stress

## WaterDemand

$$WaterDemand = K_c * PET * RIE$$

with $K_c = 1.2$, crop coefficient

## WaterSupply
```{r WaterSupply, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
WaterStressConductivity <- function(t) {sqrt(t+1) - sqrt(t)}

# Graph
t <- seq(0,30, by=0.1) 

d <- data_frame(x=t) %>%
  mutate(SoilConductivity = WaterStressConductivity(x)) %>%
  gather(Process, y, -x)

p <- ggplot(d, aes(x, y, color=Process)) + geom_line() + theme_bw() +
  labs(x="Dry Days (d)", y="SoilConductivity") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)

```

$$WaterSupply = Rainfall + Irrigation - Transpiration - Evaporation - Drainage$$

$$Transpiration = WaterDemand * WaterStressConductance$$

$$Evaporation = (1 - RIE) * PET * SoilConductivity$$

$$SoilConductivity = \sqrt{x + 1} - \sqrt{x}$$

with 

$$
  dx/dt =
  \left\{ 
	\begin{array}{ll}
	1  & \textrm{if $Rainfall + Irrigation <= 4$} \\  
	0 & \textrm{else}
	\end{array} 
	\right.
$$

## WaterStress
$$WaterStress = \frac{WaterAvailable}{WaterTotal}$$

$$WaterAvailable = WaterSupply$$

$$WaterTotal = SoilWaterCapacity * SoilDensity * (1 - StoneContent) * RootingDepth$$

## WaterStressExpansion, WaterStressConductance
```{r WaterStressProcess, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
WaterStressProcess <- function(ftsw, a=-5) {
  -1 + 2/(1 + exp(a*ftsw))
}

## Graph
ftsw <- seq(0,1, by=0.01)

d <- data_frame(x=ftsw) %>%
  mutate(
    Expansion = WaterStressProcess(x, a=-4.42),
    Conductance = WaterStressProcess(x, a=-9.3)
  ) %>%
  gather(Process, y, -x)

p <- ggplot(d, aes(x, y, color=Process)) + geom_line() + theme_bw() +
  labs(x="WaterStress (FTSW)", y="WaterStressProcess") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)
# ggsave(p, file="./figure/WaterStressProcess.png", height=4, width=4, units="in", dpi=fig_dpi)
```

$$ WaterStressProcess = -1 + \frac{2}{1 + exp^{(a * WaterStress)}}$$ 

with $a \in [-15.6;-2.3]$, genotype-dependant response parameter


## WaterStressPhenology
$$WaterStressPhenology = a * (1 - WaterStressConductance)$$
with $a=0.1$, scaling parameter for water-stress plant heating

## WaterStressMineralization
$$WaterStressMineralization = 1 - (1 - y_0) * (1 - RelativeWaterContent_{layer1})$$

\newpage

# Nitrogen stress


\newpage

# Radiation stress
## RadiationStressExpansion [@Rey2003]

```{r RadiationStressExpansion, echo=FALSE, fig.width=4, fig.height=4, fig.margin=TRUE}
# Function
RadiationStressExpansion <- function(nipar, s=2.5, a=-0.14, b=1.13, c=4.13, d=2.09) {
  s * (a + (b/(1 + exp((c - nipar)/d))))
}

## Graph
ipar <- seq(0,15,by=0.1)

d <- data_frame(x=ipar) %>%
  mutate(Expansion = RadiationStressExpansion(x)) %>%
  gather(Process, y, -x)

p <- ggplot(d, aes(x, y, color=Process)) + geom_line() + theme_bw() +
  labs(x="Intercepted Radiation per LAI (MJ)", y="RadiationStressExpansion") +
  theme(legend.direction = "horizontal", legend.position = "bottom")

print(p)
# ggsave(p, file="./figure/RadiationStressExpansion.png", height=3.5, width=4, units="in", dpi=fig_dpi)
```

$$RadiationStressExpansion = s * a + \frac{b}{1 + exp^{(\frac{c - IPAR/LAI}{d})}}$$ 

with:

* $s=2.5$, scaling parameter for density effect;
* $a=-0.14$; $b=1.13$; $c=4.13$; $d=2.09$

\newpage



# Outputs

## Timed variables

```{r OutputsTimed, echo=FALSE, results='asis'}
# data
s <- sunflo$outputs %>% select(symbol=name.rsunflo.en, label=label.doc, description=label.en, unit)
# table
kable(s, padding=0)
```

\newpage

## Indicators

```{r OutputsIndicators, echo=FALSE, results='asis'}
# data
s <- sunflo$indicators %>% select(level, factor, symbol=name.rsunflo.en, description=label.en, unit, formula)
# table
kable(s, padding=0)
```


\newpage

# References
